---
title: "Group Project - GPA Microvan Case"
author: "Team 29"
date: "2/21/2022"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r,include=FALSE}
library(readr)
#install.packages("corrplot")
library(corrplot)
#install.packages("ggplot2")
library(ggplot2)
#install.packages("cluster")
library(cluster)
#install.packages("textreg")
library(texreg)
library(dplyr)
#install.packages("Hmisc")
library(Hmisc)
#install.packages("REdaS")
library("REdaS")
#install.packages("psych")
library("psych")
#install.packages("FactoMineR")
library("FactoMineR")
#install.packages("factoextra")
library("factoextra")
#install.packages("plyr")
library(plyr)
library("ggpubr")
library(gmodels)
```

# Load Data
```{r echo = FALSE}
setwd("C:/Users/lenovo/Documents/GitHub/MI Final/")
microvan_raw <- read_csv("microvan.csv")
#View(microvan_raw)
```

# Exploratory Data Analysis
```{r}
summary(microvan_raw)
```
```{r}
str(microvan_raw, give.attr = FALSE)
num_missing <- sum(complete.cases(microvan_raw)) - nrow(microvan_raw)
print(paste("Our dataset has", num_missing, "missing values"))
```
```{r}
microvan <- microvan_raw[, -1] # Delete 1st column since it is not useful
```
## Distribution and Outliers
### Histograms
```{r}
i = 1
par(mfrow = c(5, 6), mar = c(2,2,2,2))
for (col in microvan[, -1]) {
  hist(col, main = colnames(microvan[, -1])[i])
  i = i + 1
  }
```
### Boxplots
```{r}
boxplot(x = as.list(as.data.frame(microvan[, -1])))
```
```{r}
boxplot(x = as.list(microvan[, 1:31]), las = 2)
# Removed demographics columns to better scale the 30 attributes. Results showed that lthrbetr, homlrgst, envrminr, kidsbulk, passnimp, aftrschl have outliers. So we further plot each of them.
```

```{r}
bp_lthrbetr <- boxplot(microvan$lthrbetr)
length(bp_lthrbetr$out) # 5
```
```{r}
bp_homlrgst <- boxplot(microvan$homlrgst)
length(bp_homlrgst$out) # 5
```
```{r}
bp_envrminr <- boxplot(microvan$envrminr)
length(bp_envrminr$out) # 5
```
```{r}
bp_kidsbulk <- boxplot(microvan$kidsbulk)
length(bp_kidsbulk$out) # 14
```
```{r}
bp_passnimp <- boxplot(microvan$passnimp)
length(bp_passnimp$out) # 5
```
```{r}
bp_aftrschl <- boxplot(microvan$aftrschl)
length(bp_aftrschl$out) # 5

# 5-14 outliers out of 400 sample size is the acceptable for us to conduct analysis on a 95% confidence level.
```



# Q1 Initial Model
```{r}
linear.model.1 <- lm(mvliking ~ ., data = microvan[, 1:31])
summary(linear.model.1)
#screenreg(list(linear.model.1))
```
## Confounding Variables
```{r}
microvan[, -1] %>% cor() %>% corrplot() # Visualize correlation
multicol <- microvan[, -1] %>% faraway::vif() # Multicollinearity test
print(multicol[multicol > 5]) # Problematic if greater than 5
# Turned out kidtrans/ nordtrps/ age/ miles/ numkids/ recycle are subjected to this problem
```
## Optimize Features
```{r}
linear.model.1.step <- lm(mvliking ~ ., data = microvan[, 1:31])
step(linear.model.1.step, trace = 0)
linear.model.step <- lm(formula = mvliking ~ miniboxy + lthrbetr + perfimpt + tkvacatn + 
                           noparkrm + homlrgst + suvcmpct + carefmny + shdcarpl + strngwrn + twoincom, data = microvan[, 1:31])
summary(linear.model.step)
#screenreg(linear.model.step)
```



# Q2 Factor Analysis
## Bartlett's Test of Sphericity
```{r}
bart_spher(microvan[,2:31]) # p-value < 2.22e-16, go ahead
```
## KMO-test 
```{r}
KMO(microvan[,2:31]) # 0.92 > 0.6, it is adequate.
```
## Determine the number of factors
```{r}
pca <- PCA(microvan[,2:31], scale = TRUE) # Seems to be three dimensions.
summary(pca)
pca$eig
fviz_eig(pca) # Choose 5 as factor numbers
```
## Factor solution (varimax rotation)
```{r}
fit <- factanal(microvan[,2:31], factors=5, rotation="varimax")
fit$loadings
loadings <- as.data.frame(as.table(fit$loadings))
loadings
```
## Heat Map for Factors
```{r}
seqPal5<-colorRampPalette(c("black","navyblue","mediumblue","dodgerblue3","aquamarine4","green4","yellowgreen","yellow"))(20)
as.data.frame(as.table(fit$loadings)) %>%
  ggplot(aes(Var2, Var1, fill = Freq)) +
  geom_tile()  +
  scale_fill_gradientn(name = "loadings", 
                       colors =seqPal5) +
  labs( x = "factor", y = 'attribution' )
```
## Name Factors
```{r}
scores <- factor.scores(microvan[,2:31], fit$loadings)$scores
microvan$luxury <- scores[,1]
microvan$size <- scores[,2]
microvan$kids <- scores[,3]
microvan$environment <- scores[,4]
microvan$safety <- scores[,5]
#View(microvan)
```



# Q3 Factor Regression Model
```{r}
linear.model.2 <- lm(mvliking ~ .,data = microvan[, c(1, 39:43)])
summary(linear.model.2) # f3 (kids) & f4 (environment) are insignificant 
```
```{r}
linear.model.3 <- lm(mvliking ~ luxury + size + safety ,data = microvan)
summary(linear.model.3) # All factors are significant.
```



# Q4 Cluster Anlaysis
# With all factors
## Determine the number of clusters
```{r}
d <- dist(microvan[, 39:43], method = "euclidean")
```
## Hierarchical Clustering & Dendrogram 
```{r}
hc <- hclust(d, method="ward.D") 
plot(hc, cex = 0.6, hang = -10, main = "Ward-linkage Cluster Dendrogram") # 3/ 4 might be optimal
```
## CH Index
```{r}
ch.index = function(x,kmax,iter.max=100,nstart=10,algorithm="Lloyd") {
  ch = numeric(length=kmax-1)
    n = nrow(x)
    for (k in 2:kmax) {
        a = kmeans(x,k,iter.max=iter.max,nstart=nstart,algorithm=algorithm)
        w = a$tot.withinss
        b = a$betweenss
        ch[k-1] = (b/(k-1))/(w/(n-k))
        }
    return(list(k=2:kmax,ch=ch))
}

ch_index = ch.index(microvan[, 39:43], kmax=10)
k1.max = ch_index$k[which.max(ch_index$ch)]
k1 = kmeans(microvan[, 39:43], k1.max, iter.max=100, nstart=10, algorithm="Lloyd")
plot(ch_index$k,ch_index$ch, xlab='K', ylab='CH(K)', type='b',
  main='Optimal Number of Clusters' )

# Choose 3 to be the number of clusters
```
## Cluster solution
```{r}
set.seed(42)
kmeans.solution.3 <- kmeans(microvan[, 39:43], centers = 3)
microvan$cluster3 <- kmeans.solution.3$cluster
kmeans.solution.3$centers
table(kmeans.solution.3$cluster)
```
## Heat Map for Clusters
```{r}
as.data.frame(as.table(kmeans.solution.3$centers)) %>%
  ggplot(aes(Var2, Var1, fill = Freq)) +
  geom_tile()  +
  scale_fill_gradientn(name = "center", 
                       colors =seqPal5) +
  geom_text(aes(label = round(Freq,2)), 
            colour = ifelse(as.data.frame(as.table(kmeans.solution.3$centers))$Freq > .5, "black","white" ), 
            size = 4) +
  labs( x = "factor", y = 'cluster' )
```
## Cluster Plot
```{r}
fviz_cluster(kmeans.solution.3, data=microvan[, 39:43])
```

# With only significant factors
## Determine the number of clusters
```{r}
d1 <- dist(microvan[, c(39, 40, 43)], method = "euclidean")
```
## Hierarchical Clustering & Dendrogram 
```{r}
hc1 <- hclust(d1, method="ward.D") 
plot(hc1, cex = 0.6, hang = -10, main = "Ward-linkage Cluster Dendrogram") # 3/4 might be optimal
```
## CH Index
```{r}
ch.index = function(x,kmax,iter.max=100,nstart=10,algorithm="Lloyd") {
  ch = numeric(length=kmax-1)
    n = nrow(x)
    for (k in 2:kmax) {
        a = kmeans(x,k,iter.max=iter.max,nstart=nstart,algorithm=algorithm)
        w = a$tot.withinss
        b = a$betweenss
        ch[k-1] = (b/(k-1))/(w/(n-k))
        }
    return(list(k=2:kmax,ch=ch))
}

ch_index_1 = ch.index(microvan[, c(39, 40, 43)], kmax=10)
k1.max.1 = ch_index_1$k[which.max(ch_index_1$ch)]
k2 = kmeans(microvan[, c(39, 40, 43)], k1.max.1, iter.max=100, nstart=10, algorithm="Lloyd")
plot(ch_index_1$k,ch_index_1$ch, xlab='K', ylab='CH(K)', type='b',
  main='Optimal Number of Clusters' )

# Choose 4 to be the number of clusters
```
## Cluster solution
```{r}
set.seed(42)
kmeans.solution.4 <- kmeans(microvan[, c(39, 40, 43)], centers = 4)
#microvan$cluster4 <- kmeans.solution.4$cluster
kmeans.solution.4$centers
table(kmeans.solution.4$cluster)
```
## Heat Map for Clusters
```{r}
as.data.frame(as.table(kmeans.solution.4$centers)) %>%
  ggplot(aes(Var2, Var1, fill = Freq)) +
  geom_tile()  +
  scale_fill_gradientn(name = "center", 
                       colors =seqPal5) +
  geom_text(aes(label = round(Freq,2)), 
            colour = ifelse(as.data.frame(as.table(kmeans.solution.4$centers))$Freq > .5, "black","white" ), 
            size = 4) +
  labs( x = "factor", y = 'cluster' )
```
## Cluster Plot
```{r}
fviz_cluster(kmeans.solution.4, data=microvan[, c(39, 40, 43)])
```

## Label the clusters
```{r}
microvan$segment.label <- revalue(as.character(microvan$cluster3), c("3"="lifestyle_practicality", "2"="uninvolved", "1"="affluent_size_indifferent"))
```



# Q5 Exploring Clusters
## Option 1 Cluster Regression Model 
```{r}
microvan$affluent_size_indifferent <- ifelse(microvan$segment.label == 'affluent_size_indifferent', 1, 0)
microvan$uninvolved <- ifelse(microvan$segment.label == 'uninvolved', 1, 0)
microvan$lifestyle_practicality <- ifelse(microvan$segment.label == 'lifestyle_practicality', 1, 0)
```
### Base = lifestyle_practicality
```{r}
reg.cluster1 <- lm(mvliking ~ affluent_size_indifferent+uninvolved, data = microvan[,c(1,46,47,48)])
summary(reg.cluster1)
```
### Base = uninvolved
```{r}
reg.cluster2 <- lm(mvliking ~ affluent_size_indifferent+lifestyle_practicality, data = microvan[,c(1,46,47,48)])
summary(reg.cluster2)
```
### Base = affluent_size_indifferent
```{r}
reg.cluster3 <- lm(mvliking ~ uninvolved+lifestyle_practicality, data = microvan[,c(1,46,47,48)])
summary(reg.cluster3)
```

## Option 2 T-Tests 
```{r}
group_by(microvan, microvan$segment.label) %>%
  dplyr::summarise(
    count = n(),
    mean = mean(mvliking),
    sd = sd(mvliking)
  )

ggplot(microvan) +
  geom_boxplot(aes(segment.label, mvliking, color = segment.label))+
  labs( x = "segment.label", y = "mvliking")
```
### Means between group lifestyle_practicality and group affluent_size_indifferent
```{r}
t.test_2_1 <- t.test(mvliking ~ segment.label, data = filter(microvan, segment.label != "uninvolved"))
t.test_2_1
```
### Means between group lifestyle_practicality and group uninvolved
```{r}
t.test_2_3 <- t.test(mvliking ~ segment.label, data = filter(microvan, segment.label != "affluent_size_indifferent"))
t.test_2_3
```
### Means between group affluent_size_indifferent and group uninvolved
```{r}
t.test_1_3 <- t.test(mvliking ~ segment.label, data = filter(microvan, segment.label != "lifestyle_practicality"))
t.test_1_3
```

## Option 3 Cross Tabulation
```{r}
CrossTable(microvan$segment.label, microvan$mvliking, expected=TRUE)
qchisq(0.05, 16, lower.tail = FALSE) #how to calculate the critical value = 26.29
```
### Grouping mvliking ("1-3", "4-6", "7-9")
```{r}
microvan <- microvan %>% mutate(mvliking_bin = case_when(mvliking >= 1  & mvliking <= 3 ~ '1-3',
                                                         mvliking >= 4  & mvliking <= 6 ~ '4-6',
                                                         mvliking >= 7  & mvliking <= 9 ~ '7-9'))
CrossTable(microvan$segment.label, microvan$mvliking_bin, expected=TRUE)
```



# Q6 Demographics 
```{r} 
microvan = microvan %>% 
  mutate(age_group = case_when(
    between(age, 19, 30)  ~ "19-30",
    between(age, 31, 40)  ~ "31-40",
    between(age, 41, 50)  ~ "41-50",
    between(age, 51, 60)  ~ "51-60"  )
  ) #Create age groups

microvan = microvan %>% 
  mutate(income_group = case_when(
    between(income, 15, 50)  ~ "low",
    between(income, 51, 90)  ~ "medium",
    between(income, 91, 150)  ~ "high",
    between(income, 151, 273)  ~ "very high"  )
  ) #Create income groups

microvan = microvan %>% 
  mutate(mvliking_bin = case_when(
    between(mvliking, 1, 3) ~ '1-3',
    between(mvliking, 4, 6) ~ '4-6',
    between(mvliking, 7, 9) ~ '7-9')
  ) #Create mvliking groups

microvan = microvan %>% 
  mutate(miles_group = case_when(
    between(miles, 7, 12) ~ '7-12',
    between(miles, 13, 18) ~ '13-18',
    between(miles, 19, 24) ~ '19-24',
    between(miles, 25, 32) ~ '25-32')
  ) #Create miles groups

microvan = microvan %>% 
  mutate(numkids_group = case_when(
    numkids == 0 ~ 'None',
    between(numkids, 1, 2) ~ '1-2',
    between(numkids, 3, 4) ~ '3-4')
  ) #Create numkids groups

#View(microvan)
```

```{r}
CrossTable(microvan$age_group, microvan$mvliking_bin, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```
```{r}
CrossTable(microvan$income_group, microvan$mvliking_bin, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```
```{r}
CrossTable(microvan$numkids_group, microvan$mvliking_bin, expected=TRUE)
qchisq(0.05, 4, lower.tail = FALSE)
```
```{r}
CrossTable(microvan$female, microvan$mvliking_bin, expected=TRUE)
qchisq(0.05, 2, lower.tail = FALSE)
```
```{r}
CrossTable(microvan$educ, microvan$mvliking_bin, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```

```{r}
CrossTable(microvan$miles_group, microvan$mvliking_bin, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```
```{r}
CrossTable(microvan$recycle, microvan$mvliking_bin, expected=TRUE)
qchisq(0.05, 8, lower.tail = FALSE)
```

```{r}
##Examining age-group for different clusters##
CrossTable(microvan$segment.label, microvan$age_group, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```
```{r}
##Examining income-group for different clusters##
CrossTable(microvan$segment.label, microvan$income_group, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```
```{r}
##Examining miles-group for different clusters##
CrossTable(microvan$segment.label, microvan$miles_group, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```
```{r}
##Examining numkids-group for different clusters##
CrossTable(microvan$segment.label, microvan$numkids_group, expected=TRUE)
qchisq(0.05, 4, lower.tail = FALSE)
```
```{r}
##Examining female for different clusters##
CrossTable(microvan$segment.label, microvan$female, expected=TRUE)
qchisq(0.05, 2, lower.tail = FALSE)
```
```{r}
##Examining educ for different clusters##
CrossTable(microvan$segment.label, microvan$educ, expected=TRUE)
qchisq(0.05, 6, lower.tail = FALSE)
```
```{r}
##Examining recycle for different clusters##
CrossTable(microvan$segment.label, microvan$recycle, expected=TRUE)
qchisq(0.05, 8, lower.tail = FALSE)
```